On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-05-17 Manuscript/manuscript.tex
Included New References in Introduction

\begin{abstract}
Block bootstrap is widely used in constructing confidence intervals for
parameters estimated from stationary time series. Theoretically, the method
should provide valid confidence intervals as the length of the time series goes
to infinity. In practice, however, it is necessary to know how large of a finite
sample is required for block bootstrap confidence intervals to work well. This
study aims to answer this question in a simple simulation setting where the data
are generated from a first-order autoregressive process. The empirical coverage
rates of several commonly used bootstrap confidence intervals for the mean,
standard deviation, and the lag-1 autocorrelation are compared. \jy{put most
  notable findings here. Possibly 1) large sample is needed; percentile
  bootstrap is not working well for rho (can we offer some explanation?).}

@@

Block bootstrap is a powerful tool for constructing confidence intervals (CI)
when making inferences about dependent data. Early ideas were developed
independently by \citet{hall1985resampling}, \citet{carlstein1986use}, and 
\citet{kunsch1989jackknife}.
It has since been applied in various fields such as econometrics
\citep{mackinnon2006bootstrap} and meteorology \citep{varga2017generalised}.
Block bootstrap is particularly useful for serially dependent
data when the serial dependence is not specified or not of primary interest.
The method is expected to produce CIs with coverage rates
matching their nominal levels as the sample size grows to infinity. However,
when dealing with finite sample sizes, an important question is how large the
sample size needs to be for block bootstrap CIs to have the
desired coverage rates.

The necessary sample size for bootstrap standard errors to provide valid
uncertainty measures in practice has not been extensively studied. For
independent data and non-block bootstrap, \citet{hesterberg2015teachers} notes
that while percentile-based CIs from nonparametric bootstrap
are more accurate than $t$-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. In the context of structural equation
modeling, \citet{nevitt2001performance} find
that a sample size of 200--1000 is usually sufficient for interval estimation
using standard nonparametric bootstrap. For dependent data, the necessary
sample
size for block bootstrap is even less clear. In a linear regression setting
with
dependent data, \citet{goncalves2005bootstrap} find that standard error
estimates from block bootstrap in small samples may be more accurate than
inference from closed-form asymptotic estimates, but block bootstrap percentile
\jy{did they do block or standard bootstrap?}
CIs still do not have sufficient coverage even for sample size
$n = 1024$.

\jy{I added more references to the reference folder that studied coverage of
  bootstrap CIs but not block bootstrap CIs.}

The goal of this paper is to provide some practical recommendations on
necessary sample size for block bootstrap with dependent data. We consider a
simple situation of a stationary time series, where the parameters of
interests are the mean, standard deviation, and the first-order
autocorrelation coefficient. We compare five variants of block bootstrap
CIs from the literature \citep{diciccio1996bootstrap,
  rice2006mathematical}: a basic percentile interval, a percentile
interval centered around the estimate, a bias-corrected and
accelerated ($BC_a$) interval, a standard normal interval, and a Student's $t$
interval. Their
empirical coverage rates at different sample sizes and dependence levels are
compared in a simulation study. The results of this study suggest that
appropriate recovery of temporal dependence parameters is heavily dependent on
the type of interval used.

@@

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with
$\theta$ as the mean, we just need an estimate of the standard error to
construct an approximate CI \citep[p.168]{efron1993introduction}.
Let $\widehat{\text{SE}}$ be the empirical standard error of the block bootstrap
copies of $\hat\theta_n^{(b)}$, $b = 1, \ldots, B$.
A $(1 - \alpha)$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Its validity relies on whether the distribution of $\hat\theta_n$ is reasonaly
well approximated by its asymptotic normal distribution and whether the
bootstrap $\widehat{\text{SE}}$ approximates the true standard error.

\paragraph{Student's $t$ CI}
The procedure for constructing a Student's $t$ CI is described in 
\citet{efron1993introduction}.
A $(1 - \alpha$ Student's $t$ CI is
\[
(\hat{\theta}_{n} - t_{(1-\alpha/2), k - 1}\hat{SE}, \hat{\theta}_{n} -
t_{(\alpha/2), k -1}\hat{SE}),
\]
where $k$ is the number of blocks.
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Like the standard 
normal interval, the Student's $t$ CI is a table-based bootstrap interval
approach. In this case,
we assume that $\hat\theta_n$ follows the $t_{k-1}$ distribution with an
expected value of $\theta$.

>>>>

\begin{abstract}
Block bootstrap is widely used in constructing confidence intervals for
parameters estimated from stationary time series. Theoretically, the method
should provide valid confidence intervals as the length of the time series goes
to infinity. In practice, however, it is necessary to know how large of a finite
sample is required for block bootstrap confidence intervals to work well. This
study aims to answer this question in a simple simulation setting where the data
are generated from a first-order autoregressive process. The empirical coverage
rates of several commonly used bootstrap confidence intervals for the mean,
standard deviation, and the lag-1 autocorrelation are compared. In this 
study, we find that a large sample is needed to recover even a simple
parameter like the mean, and that the selection of interval type is very
important, especially in the case of $\phi$, for which the coverage deteriorates
as the sample size increases.
\jy{put most
  notable findings here. Possibly 1) large sample is needed; percentile
  bootstrap is not working well for rho (can we offer some explanation?).}

@@

Block bootstrap is a powerful tool for constructing confidence intervals (CI)
when making inferences about dependent data. Early ideas were developed
independently by \citet{hall1985resampling, carlstein1986use, kunsch1989jackknife}.
It has since been applied in various fields such as econometrics
\citep{mackinnon2006bootstrap} and meteorology \citep{varga2017generalised}.
Block bootstrap is particularly useful for serially dependent
data when the serial dependence is not specified or not of primary interest.
The method is expected to produce CIs with coverage rates
matching their nominal levels as the sample size grows to infinity. However,
when dealing with finite sample sizes, an important question is how large the
sample size needs to be for block bootstrap CIs to have the
desired coverage rates.

The necessary sample size for bootstrap standard errors to provide valid
uncertainty measures in practice has not been extensively studied. For
independent data and non-block bootstrap, \citet{hesterberg2015teachers} notes
that while percentile-based CIs from nonparametric bootstrap
are more accurate than $t$-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. In the context of structural equation
modeling, \citet{nevitt2001performance} find
that a sample size of 200--1000 is usually sufficient for interval estimation
using standard nonparametric bootstrap. For dependent data, the necessary
sample
size for block bootstrap is even less clear. In a linear regression setting
with
dependent data where the
regression errors are generated from a homoscedastic AR(1) process, 
\citet{goncalves2005bootstrap} find that standard error
estimates from bootstrap in small samples may be more accurate than
inference from closed-form asymptotic estimates, but moving block bootstrap 
\jy{did they do block or standard bootstrap?}
\mc{moving block bootstrap}
CIs still do not have sufficient coverage of the target parameter even for sample size
$n = 1024$.
Also in the setting of linear regression, but focused on estimating the proportion of 
variation attributable to a certain variable, \citet{burch2012nonparametric} finds
that three nonparametric bootstrap methods (Standard, Percentile, and Bias-Corrected and
Accelerated) approach the coverage of a pivotal quantity as $n$ gets larger for an
underlying normal distribution. However, they find that for other distributions, 
coverage may deteriorate as $n$ increases.
\citet{chernick2009revisiting} finds that the the best choice of method for estimating
the parameter of a distribution is dependent on the sample size, the number of
bootstrap resamples, and the confidence level.
When attempting to create interval estimations for the Pearson correlation 
coefficient of bivariate normal data, \citet{puth2015variety} find bootstrap 
methods to be less accurate than the Fisher's transformation method even 
for a sample of size $n = 100$ and correlation coefficient equivalent to 0.
\jy{I added more references to the reference folder that studied coverage of
  bootstrap CIs but not block bootstrap CIs.}

The goal of this paper is to provide some practical recommendations on
necessary sample size for block bootstrap with dependent data, similar to what 
was done for basic bootstrap in \citet{hesterberg2015teachers}. We consider a
simple situation of a stationary time series, where the parameters of
interests are the mean, standard deviation, and the first-order
autocorrelation coefficient. We compare five variants of block bootstrap
CIs from the literature \citep{diciccio1996bootstrap,
  rice2006mathematical}: a basic percentile interval, a percentile
interval centered around the estimate, a bias-corrected and
accelerated ($BC_a$) interval, a standard normal interval, and a Student's $t$
interval. Their
empirical coverage rates at different sample sizes and dependence levels are
compared in a simulation study. The results of this study suggest that
appropriate recovery of temporal dependence parameters is heavily dependent on
the type of interval used.

@@

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with
$\theta$ as the mean, we just need an estimate of the standard error to
construct an approximate CI \citep[p.168]{efron1993introduction}.
Let $\widehat{\text{SE}}$ be the empirical standard error of the block bootstrap
copies of $\hat\theta_n^{(b)}$, $b = 1, \ldots, B$.
A $(1 - \alpha)$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Its validity relies on whether the distribution of $\hat\theta_n$ is reasonably
well approximated by its asymptotic normal distribution and whether the
bootstrap $\widehat{\text{SE}}$ approximates the true standard error.

\paragraph{Student's $t$ CI}
The procedure for constructing a Student's $t$ CI is described in 
\citet{efron1993introduction}.
A $(1 - \alpha$ Student's $t$ CI is
\[
(\hat{\theta}_{n} - t_{(1-\alpha/2), k - 1}\hat{SE}, \hat{\theta}_{n} -
t_{(\alpha/2), k -1}\hat{SE}),
\]
where $k$ is the number of blocks.
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Like the standard 
normal interval, the Student's $t$ CI is a table-based bootstrap interval
approach. In this case,
its validity relies on whether the distribution of $\hat\theta_n$ is
reasonably well approximation by the $t_{k-1}$ distribution with an
expected value of $\theta$, and whether the bootstrap 
$\widehat{\text{SE}}$ approximates the true standard error.
